---
title: "Loading Data"
author: "Orlando MLDS"
date: "2/8/2017"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Check to see if all packages are present on system, if not, install them.
pkgs <-c('knitr','readxl', 'jsonlite', 'XML', 'yaml')
for(p in pkgs) if(p %in% rownames(installed.packages()) == FALSE) {install.packages(p, repos='http://cran.us.r-project.org')}
for(p in pkgs) suppressPackageStartupMessages(library(p, quietly=TRUE, character.only=TRUE))
options("getSymbols.warning4.0"=FALSE)
```

# Getting Started with Data

Example code blocks showing how to load various data types into R

## .csv — Comma-Separated Values

Data type with the least overhead (aka file size) and very common

```{r}
csv_data <- read.csv("data/volcanoes.csv")
head(csv_data)
```

## .xls .xlsx — Excel Spreadsheet

Excel spreadsheets are more complicated than CSV (ex they can contain multiple datasets), but they still follow the 2D array format of a CSV and DataFrames

```{r}
require(readxl)
xlsx_data <- read_excel("data/volcanoes.xlsx") #sheet=0
head(xlsx_data)
```

## .json — JavaScript Object Notation

JSON is a common web and API communication format and easy to read

```{r}
library(jsonlite)
json_data <- fromJSON("data/volcanoes.json")
head(json_data)
```

## .xml — eXtensible Markup Language

XML looks and acts just like HTML. It is a popular web format but is being superseded by JSON

```{r}
library(XML)
xml_data <- xmlToDataFrame(xmlParse("data/volcano.xml"))
xml_data
```

## .yml .yaml — Yet Another Markup Language

YAML has less overhead than JSON and can reference other objects within the same file to save space

```{r}
library(yaml)
#yaml_data <- yaml.load_file("data/volcanoes.yml", handlers = list(map = function(x) { as.data.frame(x) }))
#head(yaml_data)
#Not yet working
```

## .rda — R Data

R Data files allow us to save and restore binary copies of native language objects

```{r}
dat <- c(1,2,3,4)
save(dat, file = "data/mydata.rda")
dat <- load(file = "data/mydata.rda")
```
